Improving Software Maintainability through Automated Refactoring of Code Clones

### 2019-10-23 saner/paper.tex
Stuff

We measure several maintainability metrics before and after refactoring to determine the impact on system maintainability of applying the refactoring. These metrics are volume, method parameters, duplication and complexity. Using this setup, we gain empirical data about the maintainability improvement of clones on a large set of data in a controlled environment. We use this data to define which characteristics of clones have the biggest influence of system maintainability when such a clone is refactored. This allows for more a accurate suggestion of code clones for refactoring and can assist in the refactoring process.

@@

\subsection{Allowing variability in a controlled set of expressions} \label{sec:t2r}
Often, duplication fragments in source code do not match exactly~\cite{kodhai2013method}. When developers duplicate fragments of code, they modify the duplicated block to fit its new location and purpose. To detect duplicate fragments with minor variance, we look into what expressions we can allow variability in, while still being refactorable.

We define the following expressions as refactorable when varied:
\begin{itemize}
  \item \textbf{Literals}: Only if all varying literals in a clone class have the same type.
  \item \textbf{Variables}: Only if all varying variables in a clone class have the same type.
  \item \textbf{Method references}: Only if the return value of referenced methods match (or are not used).
\end{itemize}
Often when allowing such variance, trade-offs come into play~\cite{krishnan2013refactoring, krishnan2014unification}. For instance, variance in literals may require the introduction of a parameter to an extracted method if the ``Extract Method'' refactoring method is used, increasing the required effort to comprehend the code.

\subsection{Gapped clones} \label{sec:t3r}
Sometimes, when fragments are duplicated, a statement is inserted or changed severely for the code to fit its new context~\cite{roy2007survey}. When dealing with such a situation, there are several opportunities to refactor so-called ``gapped clones''~\cite{ueda2002detection, zhao2018automatic}. ``Gapped clones'' are two clone instances separated by a ``gap'' of non-cloned statement(s). We define the following methods to refactor such clones:
\begin{itemize}
  \item We wrap the difference in statements in a conditionally executed block, one path for each different (group of) statement(s).
  \item We use a lambda function to pass the difference in statements from each location of the clone~\cite{tsantalis2017clone}.
\end{itemize}
For both refactoring techniques, a trade-off is at play. This is because these solutions increase the complexity and volume of the source code in favor of removing a clone.

\section{The CloneRefactor tool}\label{sec:clonerefactor}
To detect such clones that can be refactored, we surveyed a set of modern clone detection tools and techniques~\cite{svajlenko2014evaluating, sheneamer2016survey, gautam2016various, roy2009comparison}. We created a set of control projects to determine the suitability of these tools to detect refactorable clones, either through configuration or postprocessing of their output.

>>>>

We measure several maintainability metrics before and after refactoring to determine the impact on system maintainability of applying the refactoring. These metrics are volume, method parameters, duplication and complexity. Using this setup, we gain empirical data about the maintainability improvement of clones on a large set of data in a controlled environment. Because the refactoring is done automatically, we also consider micro-clones, which are shown to be important for refactoring consideration \cite{mondai2018micro}. We use this data to define which characteristics of clones have the biggest influence of system maintainability when such a clone is refactored. This allows for more a accurate suggestion of code clones for refactoring and can assist in the refactoring process.

@@

\subsection{Allowing variability in a controlled set of expressions} \label{sec:t2r}
Often, duplication fragments in source code do not match exactly~\cite{kodhai2013method}. When developers duplicate fragments of code, they modify the duplicated block to fit its new location and purpose. To detect duplicate fragments with minor variance, we look into what expressions we can allow variability in, while still being automatically refactorable.

We define the following expressions as automatically refactorable when varied:
\begin{itemize}
  \item \textbf{Literals}: Only if all varying literals in a clone class have the same type.
  \item \textbf{Variables}: Only if all varying variables in a clone class have the same type.
  \item \textbf{Method references}: Only if the return value of referenced methods match (or are not used).
\end{itemize}
Often when allowing such variance, trade-offs come into play~\cite{krishnan2013refactoring, krishnan2014unification}. For instance, variance in literals may require the introduction of a parameter to an extracted method if the ``Extract Method'' refactoring method is used, increasing the required effort to comprehend the code \cite{heitlager2007practical}.

\subsection{Gapped clones} \label{sec:t3r}
Sometimes, when fragments are duplicated, a statement is inserted or changed severely for the code to fit its new context~\cite{roy2007survey}. When dealing with such situations, there are several opportunities to refactor so-called ``gapped clones''~\cite{ueda2002detection, zhao2018automatic}. ``Gapped clones'' are two clone instances separated by a ``gap'' of non-cloned statement(s). We define the following methods to refactor such clones, based on research by Tsantalis et al~\cite{tsantalis2015assessing}:
\begin{itemize}
  \item Wrap the difference in statements in a conditionally executed block, one path for each different (group of) statement(s).
  \item Use a lambda function to pass the difference in statements from each location of the clone~\cite{tsantalis2017clone}.
\end{itemize}
For both of these techniques, a trade-off is at play. This is because these solutions increase the complexity and volume of the source code in favor of removing a clone.

\section{A tool for automatic clone refactoring}\label{sec:clonerefactor}
To detect such clones that can be refactored, we surveyed a set of modern clone detection tools and techniques~\cite{svajlenko2014evaluating, sheneamer2016survey, gautam2016various, roy2009comparison}. We created a set of control projects to determine the suitability of these tools to detect refactorable clones, either through configuration or postprocessing of their output.
